---
title: "Leporid brains"
output: html_document
---


#Loading packages
```{r load packages}

require(caper)
require(car)
require(plotly)
require(ggplot2)
library(MCMCglmm)
library(mulTree)
require(phytools)
require(forcats)
require(bayesplot)
library(hdrcde)
require(mice)
require(phylomice)
````


#Loading data and tree
```{r load data and tree}

lep.data <-read.table("./Data/lepdata_trimmed2.txt", sep = "\t", header = TRUE)
lep.tree1 <-read.tree("./Data/leptree.txt")

#convert var types

lep.data$DietBreadth <- as.factor(lep.data$DietBreadth)
lep.data$Burrow <- as.factor(lep.data$Burrow)
lep.data$Activity <- as.factor(lep.data$Activity)
lep.data$Locomotor_Mode <- as.factor(lep.data$Locomotor_Mode)

#stdize = function(x, ...) {(x - min(x, ...)) / (max(x, ...) - min(x, ...))}


lep.data$ECV_T <- log(lep.data$ECV_T)
lep.data$ECV_OB <- log(lep.data$ECV_OB)
lep.data$ECV_ROB <- log(lep.data$ECV_ROB)
lep.data$AdultBodyMass <- log(lep.data$AdultBodyMass)

#lep.data$GestationLen <- log(lep.data$GestationLen)
#lep.data$HomeRange <- log(lep.data$HomeRange)
#lep.data$LitterSize <- log(lep.data$LitterSize)
#lep.data$GR_Area <- log(lep.data$GR_Area)

#lep.data$SeP <- log(lep.data$SeP)
#lep.data$SeT <- log(lep.data$SeT)



#lep.data$ECV_T <- stdize(lep.data$ECV_T)
#lep.data$ECV_OB <- stdize(lep.data$ECV_OB)
#lep.data$ECV_ROB <- stdize(lep.data$ECV_ROB)

#lep.data$AdultBodyMass <- stdize(lep.data$AdultBodyMass)
#lep.data$GestationLen <- stdize(lep.data$GestationLen)
#lep.data$HomeRange <- stdize(lep.data$HomeRange)
#lep.data$LitterSize <- stdize(lep.data$LitterSize)
#lep.data$GR_Area <- stdize(lep.data$GR_Area)

#lep.data$SeP <- stdize(lep.data$SeP)
#lep.data$SeP <- stdize(lep.data$SeT)





#set row names and check for missingness
rownames(lep.data) <- lep.data$Name_phyl
clean.data(lep.data, lep.tree1)  ## check if data == tree names

```

#doctoring the tree
```{r fixing the tree, message=FALSE, warning=FALSE}
lep.tree1$edge.length[which(lep.tree1$edge.length == 0)] <- 0.01*median(lep.tree1$edge.length[-which(lep.tree1$edge.length == 0)])
lep.tree1 <- force.ultrametric(lep.tree1,method="extend")
```


#missingness analysis

```{r naniar}
library(naniar)
library(visdat)

#visualisation of missingness
vis_dat(lep.data)
vis_miss(lep.data)

#3% avg missing data so imputing 3 datatsets
require(sensiPhy)
lep.tree2 <- lep.tree1
lep.tree2$node.label <- NULL
phyloMiss <- miss.phylo.d(lep.data,lep.tree2, binvar=Burrow)

```

#imputation
```{r mice}
data_mice <- lep.data

library(VIM)
aggr_plot <- aggr(data_mice, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(data_mice), cex.axis=.7, gap=3, ylab=c("Histogram of missing data","Pattern"))

prec <- precomputePsi(lep.tree1)

imp <- mice(data_mice, m=3, meth = "phpmm", psi = prec$psi, psiinv = prec$psiinv, maxit = 2)

pred <- imp$predictorMatrix
pred[, "Name_phyl"] <- 0
pred[, "Se_temp"] <- 0
pred[, "Se_precip"] <- 0
pred[, "ST"] <- 0
pred[, "SP"] <- 0

imp <- mice(lep.data, m=5, pred = pred, seed=542, meth = "phpmm", psi = prec$psi, psiinv = prec$psiinv, maxit = 2000)

```



Extracting the imputed sets
```{r sets}
for(imputedsets in 1: length(imp$imp[[1]])) {
    assign(paste0("dataX",imputedsets), complete(imp, imputedsets))
    assign(paste0("dataX",imputedsets), cbind(Names = lep.data[[1]], get(paste0("dataX",imputedsets))))
}
```


Formulas
```{r formulas}

##Check all models with check_model from require(performance) as an lm first. No large vifs and
##assumptions for regression met

#Models ECV_ROB

##Seasonality
formula1a <- ECV_ROB ~ AdultBodyMass + SeT + SeP 
##Maternal investment
formula1b <- ECV_ROB ~ AdultBodyMass + LitterSize + GestationLen
##Spatial
formula1c <- ECV_ROB ~ AdultBodyMass + HomeRange + GR_Area
##Diet
formula1d <- ECV_ROB ~ AdultBodyMass + DietBreadth
##behavioural
formula1e <- ECV_ROB ~ AdultBodyMass + Locomotor_Mode + Activity + Burrow
##behavioural2
formula1f <- ECV_ROB ~ AdultBodyMass + Burrow


#Models ECT_OB

##Seasonality
formula2a <- ECV_OB ~ AdultBodyMass + SeT + SeP 
##Maternal investment
formula2b <- ECV_OB ~ AdultBodyMass + LitterSize + GestationLen
##Spatial
formula2c <- ECV_OB ~ AdultBodyMass + HomeRange + GR_Area
##Diet
formula2d <- ECV_OB ~ AdultBodyMass + DietBreadth
##behavioural
formula2e <- ECV_OB ~ AdultBodyMass +  Locomotor_Mode + Activity + Burrow
##behavioural2
formula2f <- ECV_OB ~ AdultBodyMass + Burrow





```

#Parameters for the MCMC
```{r parameters, message=FALSE, warning=FALSE}
# Number of interations
nitt <- 1000042
# Length of burnin -> what to ignore (the initial N ot iterations)
burnin <- 100000
# Amount of thinning -> lenght of step (sampling rate)
thin <- 300
```

#Creating mulTree objects
```{r multree, message=FALSE, warning=FALSE, verbose = FALSE}

for(imputedsets in 1:length(imp$imp[[1]])) {
    assign(paste("mulTree_data",imputedsets,sep=""), as.mulTree(data = get(paste("dataX",imputedsets, sep="")), tree = lep.tree1, taxa = "Name_phyl"))
}
```


PROPER PRIORS

p.var <- var(Data$BWT, na.rm = TRUE)
> prior1.1.2 <- list(G = list(G1 = list(V = matrix(p.var * 0.05),
+ nu = 1)), R = list(V = matrix(p.var * 0.95), nu = 1))

mulTree(mulTree.data = lep.data, formula = formula1f, priors = mul_priors, parameters = mul_parameters, output = ./MCMCmodels/model_1fa, ESS = 1000, chains = 2)

#change formulas
```{r multree, message=FALSE, warning=FALSE}
# The formula
# The MCMC parameters (iterations, thining, burnin)
mul_parameters <- c(nitt, thin, burnin)
# The MCMCglmm priors
mul_priors <- list(R = list(V = 1, nu = 0.002),
                   G = list(G1 = list(V = 1, nu = 0.002)))



for(imputedsets in 1 : length(imp$imp[[1]])) {
  for (form in c("2e", "1e"))
      mulTree(mulTree.data = get(paste0("mulTree_data",imputedsets)), formula = get(paste0("formula", form)), priors = mul_priors,
      parameters = mul_parameters, output = paste0("./MCMCmodels/model_", form, "/" , "model", imputedsets), ESS = 1000, chains = 2)
  }


```

``` {r pooling, message=FALSE, warning=FALSE, error = TRUE, eval = FALSE}
load("./imp.Rdata")

setwd("./MCMCmodels/model_2f")
list() -> raw_models_chain1 -> raw_models_chain2
for(imputedsets in 1:  length(imp$imp[[1]])) {
  raw_models_chain1[[imputedsets]] <- read.mulTree(paste0("model",imputedsets,"-tree1_chain1"), model = TRUE)
  raw_models_chain2[[imputedsets]] <- read.mulTree(paste0("model",imputedsets,"-tree1_chain2"), model = TRUE)
}

raw_models_allchains <- c(raw_models_chain1, raw_models_chain2)

## Get Sol
# @param model this should be a single model (e.g. one chain output from mcmcglmm)
# @return the Sol vector from one single chain
get.Sol <- function(model) {return(model$Sol)}

## Get summary of Sol
# @param model this should be a single model (e.g. one chain output from mcmcglmm)
# @return the summary table of the Sol
get.summary.Sol <- function(model) {return(summary(model$Sol))}

## Turn into dataframe and get statistics
## obtain Variance (SD^2) and bind to the list
## remove unnecessary columns
# @param The output of get.summary.Sol (the summary of Sol)
# @return dataframe with statistics
get.stat.Sol <- function(summarySol) {
  output <- as.data.frame(summarySol$statistics)
  output$Var <- (output$SD)^2
  output$SD <- NULL
  output$`Naive SE` <- NULL
  output$`Time-series SE` <- NULL
  return(output)
}

## Run the functions and get *results*
all_Sol <- lapply(raw_models_allchains, get.Sol)
all_summaries <- lapply(raw_models_allchains, get.summary.Sol)
results <- lapply(all_summaries, get.stat.Sol)

#setwd("./")

##Combining solutions for density plots + percentages

#using list.rbind from rlist
require(rlist)
solX <- as.data.frame(list.rbind(all_Sol))

##Neat up the names!

names(solX)
names(solX) <- c("Intercept", "Body Weight", "Burrow")



#Plot density plots

#plot in pdf


pdf(file="./density.pdf")
par( mfrow = c( 2, 2 ))
    for(column in 1 : c(ncol(solX))) {
    percentage_above_zero <- length(which(solX[,column] >= 0))/length(solX[,column])
hdr.den(solX[,column],  main = names(solX[column]), sub = paste0(round(percentage_above_zero*100,2), "% above zero"), prob = c(50, 95, 99))
abline(v = 0, lty = 1)  
    }
dev.off()
```






##Pooling the results using the Rubin's rule
##Make sure to adjust parameters!
```{r pooling, message=FALSE, warning=FALSE, error = TRUE, eval = FALSE}

# Using the posteriors collated in *results*

mbar <-  function (x, col=1) { # function to calculate average parameter estimates and average variances from the results list
    vals <- rowMeans(matrix(unlist(lapply(x, function (z) z[, col])),
                            nrow=dim(x[[1]])[1], ncol=length(x)))
    names(vals) <- rownames(x[[1]])
    vals
}

Bm <- function (x) {
    Qmbar <- mbar(x)
    Qvals <- matrix(unlist(lapply(x, function (z) z[,1])),
                    ncol=length(x), nrow=dim(x[[1]])[1])
    QQ <- apply(Qvals, 2, function (x) x - Qmbar)
    apply(QQ, 1, function (z) (z %*% z)/(length(z)-1))
}

Tm <- function (x) {
    Umbar <- mbar(x,  col=2)
    Umbar + (1 + 1/length(x)) * Bm(x)
}

lambda <- function (nu) {
    (nu + 1)/(nu + 3)
}

vm <- function (Bm, Tm, m=5){
    gammahat <- (1+1/m) * sum(Bm/Tm)/length(Bm)
    (m-1)/(gammahat^2)
}


# Now we do the analysis:


#m - imputed sets * chains
#n - number of species
#k - number of params as N of cols in solutions

m <- (length(imp$imp[[1]]))*2
n <- 18
k <- ncol(solX)
## m = number of imputed datasets
## n = number of observations
## k = number of parameters

Bm1 <- Bm(results)
Tm1 <- Tm(results) # total variance

# calculation of the degrees of freedom for t-tests of parameters

vhatobs <- lambda(n-k)*(n-k)*(1- (1+1/m) * sum(Bm1/Tm1)/length(Bm1))
vm1 <- vm(Bm1, Tm1)
vmtilde <-  1/(1/vm1+1/vhatobs)

Qmbar <- mbar(results) # mean parameter estimates

WaldT <- Qmbar/sqrt(Tm1)
upperCI <- Qmbar + sqrt(Tm1) * qt(.95, vmtilde)
lowerCI <-  Qmbar - sqrt(Tm1) * qt(.95, vmtilde)

tTable <- cbind(Qmbar, SE=sqrt(Tm1), WaldT,
                df=vmtilde, p=2*(1-pt(abs(WaldT),
                                    vmtilde)), lowerCI, upperCI)
```

##DIC calculation
```{R comparing DIC, error = TRUE, eval = FALSE}
#Getting DICs and averaging (possibly other ICs)

get.DIC <- function(model) {return(model$DIC)}

model_DICs <- lapply(raw_models_allchains, get.DIC)  
model_DICs <- as.data.frame(model_DICs)
DICnames <- c(paste0("Model", 1:ncol(model_DICs)))
colnames(model_DICs) <- DICnames
DICs <- mean(t(model_DICs))
DICs
```

##H calculation
```{r h - phylogenetic signal, error = TRUE, eval = FALSE}
#Calculating average H for all models

get.H <- function(model) {return((var(model$VCV[,"animal"]))/
+     (var(model$VCV[,"animal"]) + var(model$VCV[,"units"])))
  }

Hs <- lapply(raw_models_allchains, get.H)
Hs <- as.data.frame(Hs)
mean(t(Hs))

```

##Plotting 2: Export text and plot output
```{r sink to txt file, error = TRUE, eval = FALSE}
#Change row names
row.names(tTable)
row.names(tTable) <- names(solX)

#Export a csv table
#col.names = NA offsets the header with 1
write.table(as.matrix(tTable), "./analysis-output.csv", sep = ",", col.names = NA, row.names = TRUE)

#Export txt with posteriors and Hs and DICs
sink('./analysis-output.txt')
#options(width=10000) stops word wrapping
options(width=10000)
#print pooled results
tTable
#print Hs
print("Mean H")
mean(t(Hs))
#print DICs
print("Mean DIC")
DICs
sink()

#Convert data to res.plot
res.plot <- as.data.frame(tTable)

#res.plot$upperCI <- NULL
#res.plot$lowerCI <- NULL
res.plot$p <- NULL
res.plot$df <- NULL
res.plot$WaldT <- NULL

#generate pdf with the model
pdf(file="./model.pdf",width=6,height=4)
p <- ggplot(
    res.plot, 
    aes(x = Qmbar, y = fct_relevel(row.names(res.plot), "VarN", "Intercept", after = Inf), xmin = lowerCI, xmax = upperCI)) +
    geom_point(aes(color = row.names(res.plot))) +
    geom_errorbarh(aes(color = row.names(res.plot)), height=0.05)+
    geom_vline(xintercept = 0, color = "red", linetype="longdash", size=0.5) +
    theme_light()
p$labels$colour <- "Model parameters"
p$labels$x <- "Posterior estimate + 95% CI"
p$labels$y <- "Model parameters"
p

#generates bayesplot of the model
color_scheme_set("brightblue")
mcmc_intervals(t(res.plot))

dev.off()

#optional
mcmc_areas(t(res.plot))
```




##PGLS and lambda
```{r pgls, error = TRUE, eval = FALSE}


pgls_m1 <- gls(ECV_ROB ~ AdultBodyMass + SeT + SeP, data = lep.data, correlation = corPagel(1,lep.tree1, form = ~Name_phyl), na.action = na.omit)
anova(pgls_m1)
summary(pgls_m1)
lm_m1 <- lm(ECV_ROB ~ AdultBodyMass * Burrow, lep.data)
check_model(lm_m1)


formula1a <- ECV_ROB ~ AdultBodyMass + SeT + SeP 
##Maternal investment
formula1b <- ECV_ROB ~ AdultBodyMass + LitterSize + GestationLen
##Spatial
formula1c <- ECV_ROB ~ AdultBodyMass + HomeRange + GR_Area
##Diet
formula1d <- ECV_ROB ~ AdultBodyMass + DietBreadth
##behavioural
formula1e <- ECV_ROB ~ AdultBodyMass + Locomotor_Mode + Activity + Burrow
##behavioural2
formula1f <- ECV_ROB ~ AdultBodyMass + Burrow


pgls_m1 <- gls(ECV_ROB ~ AdultBodyMass + Burrow, data = lep.data, correlation = corPagel(1,lep.tree1, form = ~Name_phyl), na.action = na.omit)

lambda <- seq(-1.6, 1, length.out = 500)

lik <- sapply(lambda, function(lambda) logLik(gls(ECV_ROB ~ AdultBodyMass + Burrow,correlation = corPagel(form=~Name_phyl, value = lambda, phy = lep.tree1, fixed = TRUE) ,data = lep.data, na.action = na.omit)))

plot(lik ~ lambda, type = "l", main = expression(paste("Likelihood Plot for ",lambda)), ylab = "Log Likelihood", xlab = expression(lambda), xlim=c(-0.5, 1), ylim=c(-1, 1))

abline(v = pgls_m1$modelStruct, col = "red")


rect(0, -1, 1, 1, col= rgb(0,0.4,0,alpha=0.2), border = NA)
text(-0.1840879, 0.5248559, "-0.33")
```


##Graph per species
```{r old pgsl}
require (ggConvexHull)

colourCount = nlevels(individuals$Species)
getPalette = colorRampPalette(brewer.pal(8, "Accent"))
legend_ord <- levels(with(individuals, reorder(Species, Mass)))

ggplot(individuals, aes(log(Mass), log(ECV), col=Species, shape=Burrow)) +
    geom_point(size=2) +
    geom_smooth(method=lm, se=FALSE, fullrange=FALSE, size=1.3, aes(colour=factor(Burrow))) +
    geom_convexhull(alpha = 0.05,aes(color=Burrow), show.legend = F) +
    stat_summary(fun="mean", colour="red", geom="point", size=4, shape=4) +
    xlab("log Body weight") + ylab("log Brain volume") +
    scale_color_discrete (breaks=legend_ord) +
    scale_fill_manual(values = getPalette(colourCount), name = "Species") +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
    panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    guides(fill = guide_legend(override.aes = list(linetype = 0)),
         color = guide_legend(order =1, override.aes = list(linetype = 0)))
```